---
title: "Tutorial: Install Fest"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

This tutorial will introduce you to the main softwares we will be using throughout our workshop. You need to run this tutorial to have everything in place to fully participated in our classes. 

# R

We will use **R** as the software and programming language in this course. R is a free programming language, extremely flexible and with an excellent support community. It is now one of the most used software by social scientists and data scientists in general. For this reason, I chose to use R.

A super frequently asked question when I teach other iterations of this workshop is: **Why  R?**. The answer for me is simple: R is unbeatable when it comes to manipulating databases (the rectangular Excel-style datasets we love), data visualization and statistical modeling. In addition, R is also a programming language, which allows you to do everything that other more general languages do, together with the excelent features with regards to data manipulation, visualization, and modeling tools mentioned above.

Below you can see how to install R on your local machine. 

### Windows

To install R on Windows, the first step is to download the program. To do this, enter [this link ](https://cran.r-project.org/bin/windows/base/) and click on “Download R 4.03 for Windows”.

You will download it, and the file will probably be saved in your Downloads folder. Go to this folder, double click on the file and start installing **R**.

### Mac OS

To install R on Mac OS, the first step is to download the program. To do this, enter [this link ](https://cran.r-project.org/bin/macosx/) and click on “Download R 4.03”.

# Rstudio


**R**  is a programming language that, after you install it, you can access on your computer on many different ways. You can call **R** from your command line, as old-school programmers usually do. Or you can use other softwar and programs developed to make the usage of the language more intuitive. These softwares are called IDEs -- integrated development environment. 

The most famous and useful IDE for R is called **RStudio**. Without goint
O IDE mais utilizado em R chama-se **RStudio**. Without delving into computer science concepts too much, R Studio consists of a program that integrates R with other tools, and makes using R more intuitive. For example, R Studio is already integrated with other programs (like git), allows multiple tasks to be done with point-and-click, allows **auto-completion** of programming functions and indicates when there are errors in your code.

To download RStudio, you can click [here](https://rstudio.com/products/rstudio/download/). You should choose the version corresponding to your operating system.

Once installed, to use R, you can go directly to R Studio, and open it. Then we're ready to start programming.


## R Studio Tutorial 

## Tools --> Global Options

Before we begin, let's set a few RStudio settings to improve your experience.  

1. Click "Tools --> Global Options --> Appearance" to change your color and font settings.  

2. Click "Tools --> Global Options --> Code" and check the box that says "Soft-wrap R source files" to wrap the text in your script to the width of the script pane.  

3. Click "Tools --> Global Options --> Code --> Display" and check the boxes that say "Highlight selected line" and "Highlight R function calls". 

# The basics: Navigating RStudio

Open RStudio! Then, open a new script by clicking "File --> New File --> R Script" or by pressing `Ctrl + Shift + N` (PC) or `command + shift + N` (Mac). After opening a new script, you should see four window "panes".  

## Top left pane (input/script)

Enter code in this savable `script` file in the top left pane. This is a plain text file but with a .R extension. Enter `2 + 2` in your script and run a line of code by pressing `command + enter` (Mac) or `Ctrl + enter` (PC). Or, click the "Run" button at the top of the script.  

A hashtag `#` tells R that you do not want that particular line or block of code to be run - this is called **commenting your code**. This is handy for making notes to yourself and you can even add hashtags after lines of runable code, on the same line.  

The name of your script file is in the tab at the top of your script window - the name defaults to `Untitled1`. Be sure to save your script by clicking "File --> Save" or `command + s` (Mac) or `Ctrl + s` (PC). You can also click the floppy disk icon to save. 

## Bottom left pane (output/console)

Code output is displayed in the `console` in the bottom left pane. This space is also good for just noodling around and trying out code that you do not wish to save in your script.  

In the console, the prompt `>` looks like a greater than symbol. If your prompt begins to look like a `+` symbol by mistake, simply click in your console and press the `esc` key on your keyboard as many times as necessary to return to the prompt. 

R uses `+` when code is broken up across multiple lines and R is still expecting more code. A line of code does not usually end until R finds an appropriate stop parameter or punctuation that completes some code such as a closed round parenthesis `)`, square bracket `]`, curly brace `}`, or quotation mark `'`.  

If the output in your console gets too messy, you can clear it by pressing `control + l` on both Mac and PC. This will not erase any saved data - it will simply make your console easier to read. 

## Top right pane (global environment)

Data are saved in R's memory as "variables". Variables are simply placeholders for a value, mathematical expression, word, function, or dataset! The global "Environment" tab in the upper right pane displays the variables you have assigned/saved. "Global" simply means that these variables are available for any task.  

## Bottom right pane (files, plots, packages, and help)

Here you find useful tabs for navigating your file system, displaying plots, installing packages, and viewing help pages. Press the `control` key and a number (1 through 9) on your keyboard to shortcut between these panes and tabs. 

### <span style='color:#e13d3d'> EXTRA: Instalando pacotes no R </span> 


There are a number of `packages` that are supplied with the R distribution. These are known as ``[base packages](https://stat.ethz.ch/R-manual/R-devel/library/base/html/00Index.html)" and they are in the background the second one starts a session in R.

Packages are collections of R functions, data, and compiled code in a well-defined format. To use a package you need to go through the following steps:



1. Install the package -- **only once**. 
2. Activate the package -- **every time you open a new R session**. 

On several occasions you will see conflicts appearing between packages, or errors from packages with other dependencies in your machine. That's why I added this extra section where you can install some very useful packages in your R. These are the main packages we will use in this workshop.

Copy and paste these lines below and run (command + enter` (Mac) or `Ctrl + enter` (PC)) each line to install the packages.


```{r, eval=FALSE}
# Instalando um pacote. 

```{r, eval=FALSE}
# Instalando um pacote.
install.packages("tidyverse") # conjunto de pacotes para manipulação e visualizaco
install.packages("quanteda") # análise de texto
install.packages("tidytext") # análise de texto
install,packages("rebus") # análise de texto
install.packages("ggthemes") # Visualização
install.packages("ggrepel") # Visualizacao
install.packages("here") # código reproduzível
install.packages("stm") # modelo de topics
install.packages("tidystm") # modelo de topicos
install.packages("RColorBrewer") # Cores bonitas
install.packages("tidymodels") # modelagem
install.packages("estimatr") # modelagem
install.packages("rjson") #manipular json
install.packages("rtweet") # twitter
install.packages("rvest") #scrapping
install.packages("httr") #scrapping
install.packages("xml2") # scrapping
install.packages("electionsBR") # dados eleitorais no Brasil
```


### <span style='color:#e13d3d'> EXTRA II: Any error? </span> 

Do not worry! Errors in installing packages are very common. And among the most useful elements of using R is the large number of resources available online and an excellent community of programmers willing to help you going through these problems.

First, if you have problems installing packages, make sure you are connected to the internet and that https://cloud.r-project.org/ is not blocked by your firewall or proxy.

Second, look for help on google. I've been programming in R for almost ten years, and I cannot programm without asking for help on google. Asking for help online is simple: copy and paste R's message into google. And go get the solutions. The main solutions site is called **stackoverlflow**. Most of the solutions there will solve your problems. 

## RMarkdown

Several of the materials I will use in the course were generated using **Rmarkdown**.  The **R Markdown** is a packages that allows you to combine text (markdown) with code (R) and generate high quality documents in html, doc and pdf. You can read more [here](https://bookdown.org/yihui/rmarkdown/)

This combination between text and code sounds simple, but it is not.  It is a complex and extremely useful  capability. For example, this tutorial, as well as my slides, were all done using Rmarkdown. We will explore together the potential of Rmarkdown in the next classes. For now, let's just install the package.

Open an R script on your laptop and run the following code.

```{r eval=FALSE}
# Install from CRAN
install.packages('rmarkdown')
```

### Rmarkdown to pdfs.

To produce documents using Rmarkdown in PDF format, we need to have LaTeX installed in our operating system. Latex is a text markup language -- a syntax in which you write text with layout indications, as it should appear on the final page. In other words, you don't immediately see the output of your writing.  Instead of using word point-and-click, LaTeX uses a series of functions to produce its layouts, which, consequently, makes it more flexible, and very useful for academic text production.

For R Markdown users who are not LaTeX users, it is suggested that you install TinyTeX (https://yihui.name/tinytex/).

```{r eval=FALSE}
install.packages('tinytex')
tinytex::install_tinytex()  # install TinyTeX
```

## Google Chrome

You will need to have a modern web browser installed to perform some of the tasks in this lesson. The recommended browser for this class is [Google Chrome](https://www.google.co.uk/intl/en_uk/chrome/).

## Gadget Selector.

As part of the data scraping classes, you will need to examine the HTML elements in your data. In this lesson, we'll use the Selector Gadget for this purpose. This tool allows us to easily examine the elements behind each webpage.

If you are using Google Chrome, you can simply install the [Selector Gadget extension in Chrome](https://chrome.google.com/webstore/detail/selectorgadget/mhjhnkcfbdhnjickkkdbjoemdmbfginb?hl=en). If for some reason you can't use Chrome extensions (including in Chrome itself), you can install Selector Gadget by following the instructions on the Selector Gadget website.

## Conclusion.

For today, we're done. 

## See you!!!