---
title: "Twitter API"
subtitle: "IPSA-Flacso Summer School"
author: "Tiago Ventura"
institute: ""
date: ""
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r setup, include=FALSE}

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(messagwese=FALSE, warning = FALSE)
xaringanthemer::style_mono_accent(base_color ="#23395b", 
                                  title_slide_text_color="#23395b", 
                                  title_slide_background_color = "#fbf9f4", 
                                  background_color = "#fbf9f4", 
                                  link_color =  "#F21A00", 
                                 code_font_size = "60%")

options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(message=FALSE, warning = FALSE, error=TRUE)


``` 


# Introduction

Today, we will do a brief overview about using the Twitter Rest and Stream Apis.

---

# Accessing Twitter Data. 

To access the Twitter API, we will use the [rtweet](https://github.com/ropensci/rtweet) package. Our first step is to install the package.
```{r}

#devtools::install_github("ropensci/rtweet", force=TRUE)
#install.packages("rtweet") # somente uma vez
library(rtweet)
library(tidyverse)
library(broom)
```

---

## Getting your Credentials

As we learned in the last class, some APIs require prior registration to give proper authorization to access their data. 

This is the case with Twitter. To request credentials, you need to:

- have a twitter account,

- create a twitter developer account. 

The `rtweet` package has an excellent tutorial on how to request developer access.

```{r, eval=FALSE}
vignette("auth", package = "rtweet")
```

---

After you get your credentials, you have to tell them to R. We do this creating objects (this is actually a functionality from the rtweet package)

```{r eval=FALSE}
app_name<-"Tiago Ventura"

consumer_key="X"
consumer_secret="X"
access_token<- "X"
access_token_secret<-"X"
```

Create your accesss

```{r}
create_token(app=app_name,
             consumer_key=consumer_key, 
             consumer_secret=consumer_secret,
             access_token = access_token, 
             access_secret = access_token_secret)


```

This functions saves your credentials in your environment. You don't need to repeat this in the future. 

---

## Rest API: Work with to get old tweets. 

Twitter has two APIs: Rest API and Stream API. Let's start with the REST API. This API allows:

- Access to tweets from the last 6-9 days.

- 18,000 tweets in every 15-minute hit.

To use this function, you need to enter a search term. Twitter's advanced search helps you format the appropriate search terms when your interest is in more than a single word.

```{r}
bolsonaro_tweets<-search_tweets("bolsonaro", n=50, include_rts = TRUE)
bolsonaro_tweets %>% as_tibble()
```

---

## get the timelines

Allows you to get the timeline of certain profiles
```{r}
timelines_covid <- get_timelines(c("renancalheiros", "ottoalencar", "EduGiraoOficial"), n=100)
timelines_covid %>% as_tibble()
```

---

## Information about the users


```{r}

users <- lookup_users(c("renancalheiros", "ottoalencar", "EduGiraoOficial"))

users %>%as_tibble()
```

---

## Recent likes
```{r}
favs <- get_favorites(c("renancalheiros", "ottoalencar", "EduGiraoOficial"))
favs %>% as_tibble()
```

---
## And followers

```{r}

follow_renan <- get_followers("renancalheiros")
follow_renan %>% as_tibble()
```

---
# API Stream

Twitter has a second API where you can sample the tweets being produced in real time. This API gives you more access to data, so it's the best way to collect it, you can leave a few hours -- or days -- running on your R.

```{r}
election <- stream_tweets("", 
                          timeout = 100, 
                          file_name = "file.json")
election

```

---

That is it. 

Next week you will learn a lot of different ways to analyze all this data!


---

## Academic API

Twitter has recently launched a new api, called academic API, that gives you access to historical data on Twitter. It is amazing. 

To use it, you need a different authorization. And also another r package. Here is some places where you can find the information. 

[a 101 course](https://github.com/twitterdev/getting-started-with-the-twitter-api-v2-for-academic-research)

[academictwitteR](https://github.com/cjbarrie/academictwitteR)
